---
title: "Spatial Sampling, then OLS"
output: html_document
---

1. Sample cells at 100km distance to skip spatial autocorrelation
2. Regular OLS
3. Model evaluation.
4. Repeat above for consistency

## read data

```{r}

library(ggplot2)
library(tidyverse)
library(broom)
library(dplyr)
library(rgdal)
library(rgeos)
library(raster)
library(sf)
library(sp)
library(tmap)
library(viridis)
library(spdep)
library(spatialreg)
library(lmtest)
library(Metrics) # for rmse
library(leaflet)

#E:/R-code/Modeling/code/FCN_clean_csvs.R
#~/Documents/R-code
source('E:/R-code/Modeling/code/FCN_clean_csvs.R')

MT_outline <- readOGR(dsn = 'E:/R-code/Modeling/data/shp/MatoGrossoOutline', layer = 'MatoGrossoOutline')
crs(MT_outline) <- CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")

median_cell_raw <- read.csv('E:/R-code/Modeling/data/median_onset_cell_v2.csv')

grid_1deg <- readOGR(dsn = 'E:/R-code/Modeling/data/shp/grid_1deg', layer = 'grid_1deg')
munis <- readOGR(dsn = 'E:/R-code/Modeling/data/shp/munis', layer = 'munis_SHP')
crs(munis) <- CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")
  
cell_sf <- st_read(dsn = 'E:/R-code/Modeling/data/shp/median_onset_cell', layer = 'median_onset_cell_SHP')

min_soy_area <- 2 #km2. min area of total or SC/DC soy in cell, muni or property to be considered in model

```

## functions to evaluate model and make plots

```{r}
# test the models
test_plots <- function(model, model_name) {
  
  plot(model$residuals, main = paste("residual vs index", model_name))

  plot(model$fitted.values, model$residuals, main = paste("fitted values vs residuals", model_name))
  
  # normal probability plot
  qqnorm(model$residuals, pch = 1, frame = FALSE, main = model_name)
  qqline(model$residuals, col = "steelblue", lwd = 2)
}

# add cell_ID
clean_cell_ID <- function(cell_ID) {
  strsplit(cell_ID, "_")[[1]][2]
}

# plot data
plot_cell_onset <- function(year, cell_data) {
  cell_year <- cell_data[cell_data$year == year, ]
  
  ggplot(cell_year) +
   geom_sf(aes(fill = onset)) +
   scale_fill_viridis() +
   ggtitle(paste("Onset for spatial sampling", year)) +
   geom_polygon(data = MT_outline, aes(x = long, y = lat), color = "black", alpha = 0, linetype = 1) +
   theme_bw()
}

plot_cell_residuals <- function(year, cell_data, intensity) {
  cell_year <- cell_data[cell_data$year == year, ]
  
  ggplot(cell_year) +
   geom_sf(aes(fill = residuals)) +
   scale_fill_viridis() +
   ggtitle(paste(intensity, "residuals for spatial sampling", year)) +
   geom_polygon(data = MT_outline, aes(x = long, y = lat), color = "black", alpha = 0, linetype = 1) +
   theme_bw()
}

plot_cell_plant <- function(year, cell_data, intensity) {
  cell_year <- cell_data[cell_data$year == year, ]
  
  ggplot(cell_year) +
   geom_sf(aes(fill = plant)) +
   scale_fill_viridis() +
   ggtitle(paste(intensity, "plant for spatial sampling", year)) +
   geom_polygon(data = MT_outline, aes(x = long, y = lat), color = "black", alpha = 0, linetype = 1) +
   theme_bw()
}

plot_cell_tempAuto <- function(year, cell_data, intensity) {
  cell_year <- cell_data[cell_data$year == year, ]
  
  ggplot(cell_year) +
   geom_sf(aes(fill = dwi.autocorr.p.value)) +
   scale_fill_viridis() +
   ggtitle(paste(intensity, "signif temporal autocorrelation for spatial panel", year)) +
   geom_polygon(data = MT_outline, aes(x = long, y = lat), color = "black", alpha = 0, linetype = 1) +
   theme_bw()
}


```

## read and clean spatial data

```{r}

# CSV DATA -----------------------------------------------------------------------------------------------------------------
# median cell
median_cell <- median_cell_raw %>% delete_cols_median_cell() %>%
                                    rename_cols_median_cell()

# SF DATA ------------------------------------------------------------------------------------------------

cell_sf$cell_ID <- median_cell$cell_ID
cell_sf$cell_ID <- sapply(as.character(cell_sf$cell_ID), clean_cell_ID)


cell_sf_tidy <- cell_sf %>% tidy_by_intensity_plant("SC_plant", "DC_plant") %>%
            #tidy_by_intensity_delay("SC_delay", "DC_delay") %>%
            dplyr::select(-c(SC_harvest, DC_harvest))
cell_sf_tidy$year_index <- cell_sf_tidy$year - 2003
cell_sf_tidy$year_factor <- as.factor(cell_sf_tidy$year)

cell_sf_tidy <- cell_sf_tidy %>%  drop_na

```

## functions to run model

```{r}

get_sampled_data <- function(full_data, grid_size, lat_offset, lon_offset, agg_scheme, plot_samples, year_oi) {
  # agg_scheme is FALSE for using only individual cells at grid point locations, TRUE for aggregating to grid polygons by mean cell values. plot_samples is TRUE or FALSE, for showing where samples were taken for SC and DC for a given year_oi. year_oi is only for plotting purposes
  
  # order cells so the slowest-changing row is cell_ID
  full_data <- full_data[order(full_data$cell_ID) , ]
  full_data$index <- 1:nrow(full_data) #unique row index to join later
  full_data_sp <- as(full_data, 'Spatial')
  
  # generate regular grid of points
  samplePoints <- makegrid(MT_outline, cellsize = grid_size)
  samplePoints$x1 <- samplePoints$x1 + lon_offset
  samplePoints$x2 <- samplePoints$x2 + lat_offset
  samplePoints <- SpatialPoints(samplePoints, proj4string = CRS(proj4string(MT_outline)))

  # sampling only the desired 
  if(!agg_scheme) {
    # sample from grid of points, delete NA's
    samples <- sp::over(samplePoints, full_data_sp, returnList = TRUE) %>%
                bind_rows()
    samples <- samples[complete.cases(samples),] # data.frame
    sampled_data <- merge(data.frame(index = samples[, "index"]), full_data, by.x = "index", by.y = "index")
    sampled_data <- st_as_sf(sampled_data)
  }
  
  if(agg_scheme) {
    samplePixels <- SpatialPixels(samplePoints[MT_outline,])
    samplePixels <- as(samplePixels, "SpatialPolygons")
    
    # turn SpatialPolygon into SpatialPolygonsDataFrame
    # Create a dataframe and display default rownames
    samplePixels.df <- data.frame( ID=1:length(samplePixels))
    rownames(samplePixels.df)
        
    # Extract polygon ID's
    pid <- sapply(slot(samplePixels, "polygons"), function(x) slot(x, "ID"))
        
    # Create dataframe with correct rownames
    samplePixels.df <- data.frame( ID=1:length(samplePixels), row.names = pid)  
        
    #coersion 
    samplePixels.spdf <- SpatialPolygonsDataFrame(samplePixels, samplePixels.df)
    
    # separate full_data_sp by intensity because can't take mean of intensity. also get rid of cell_ID, year_factor
    full_data_sp_DC <- full_data_sp[full_data_sp$intensity == "DC",
                                    !(names(full_data_sp) %in% c("cell_ID", "intensity", "year_factor"))]
    full_data_sp_SC <- full_data_sp[full_data_sp$intensity == "SC",
                                    !(names(full_data_sp) %in% c("cell_ID", "intensity", "year_factor"))]
    
    # 'empty' SpatialPolysDataFrame to store per-year info
    sampled_data_DC <- full_data_sp_DC[1,]
    sampled_data_SC <- full_data_sp_SC[1,]
    
    # aggregate 25km cells to grid cell. do by year and intensity separately
    for (year in 2004:2014) {
      sampled_year_DC <- aggregate(full_data_sp_DC[full_data_sp_DC$year == year,], by = samplePixels.spdf, FUN = mean) 
      sampled_year_SC <- aggregate(full_data_sp_SC[full_data_sp_SC$year == year,], by = samplePixels.spdf, FUN = mean) 
      
      sampled_data_DC <- rbind(sampled_data_DC, sampled_year_DC)
      sampled_data_SC <- rbind(sampled_data_SC, sampled_year_SC)
    }
    
    # get rid of first row
    sampled_data_DC <- sampled_data_DC[2:nrow(sampled_data_DC),]
    sampled_data_SC <- sampled_data_SC[2:nrow(sampled_data_SC),]
    
    # add intensity, year_factor as a new row
    sampled_data_DC$intensity <- rep("DC", nrow(sampled_data_DC))
    sampled_data_SC$intensity <- rep("SC", nrow(sampled_data_SC))
    sampled_data_DC$year_factor <- as.factor(sampled_data_DC$year)
    sampled_data_SC$year_factor <- as.factor(sampled_data_SC$year)
    
    # combine SC and DC
    sampled_data <- rbind(sampled_data_DC, sampled_data_SC)
    sampled_data <- st_as_sf(sampled_data)
    
    sampled_data <- sampled_data %>% drop_na

  }
  
  if (plot_samples & agg_scheme) {
    sample_year <- subset(sampled_data, year == year_oi)
    full_year <- subset(full_data, year == year_oi)
    
    DC_sampled_map <- ggplot() +
               geom_sf(data = subset(full_year, intensity == "DC"), color = "gray") +
               geom_sf(data = subset(sample_year, intensity == "DC"), color = "blue", alpha = 0) +
               ggtitle(paste("Sampled cells, for DC", year_oi)) +
               geom_polygon(data = MT_outline, aes(x = long, y = lat), color = "black", alpha = 0, linetype = 1) +
               theme_bw()
    
    SC_sampled_map <- ggplot() +
               geom_sf(data = subset(full_year, intensity == "SC"), color = "gray") +
               geom_sf(data = subset(sample_year, intensity == "SC"), color = "blue", alpha = 0) +
               ggtitle(paste("Sampled cells, for DC", year_oi)) +
               geom_polygon(data = MT_outline, aes(x = long, y = lat), color = "black", alpha = 0, linetype = 1) +
               theme_bw()
        
    print(DC_sampled_map)
    print(SC_sampled_map)
  }
  
  if (plot_samples & !agg_scheme) {
    # plot the samples over the full dataset for a given year
    full_year <- subset(full_data, year == year_oi)
    sample_year <- subset(sampled_data, year == year_oi)
    
    DC_sampled_map <- ggplot() +
           geom_sf(data = subset(full_year, intensity == "DC"), color = "gray") +
           geom_sf(data = subset(sample_year, intensity == "DC"), color = "blue") +
           ggtitle(paste("Sampled grid, for DC", year_oi)) +
           geom_polygon(data = MT_outline, aes(x = long, y = lat), color = "black", alpha = 0, linetype = 1) +
           theme_bw()
    
    SC_sampled_map <- ggplot() +
           geom_sf(data = subset(full_year, intensity == "SC"), color = "gray") +
           geom_sf(data = subset(sample_year, intensity == "SC"), color = "blue") +
           ggtitle(paste("Sampled grid, for SC", year_oi)) +
           geom_polygon(data = MT_outline, aes(x = long, y = lat), color = "black", alpha = 0, linetype = 1) +
           theme_bw()
    
    print(DC_sampled_map)
    print(SC_sampled_map)
  }
  
  # output: data.frame to input to models
  return(sampled_data)
  
}

# test output
#sampled_data <- get_sampled_data(cell_sf_tidy, grid_size = 1, lat_offset = 0.5, lon_offset = 0.5, agg_scheme = TRUE, plot_samples = TRUE, year_oi = 2005)
```

## functions to evaluate different sampling strategies and predictors

```{r}

# cross validation by location
run_model_for_prediction_elimlocation <- function(full_data, formula, year_for_location, year_oi, plot_model_evals) {
  
  # year_oi is for mapping and visualization only
  # year_for_location is the year whose locations are looped through in cross validation
  
  # use only data points that existed in one year (2004)
  data_in_year <- full_data[full_data$year == year_for_location,]
    
  prediction_results <- data.frame()
  
  for (row_index in 1:nrow(data_in_year)) {
    test_location_sf <- data_in_year[row_index,]
    
    # turn test location into point, otherwise will get multiple cells as test location
    test_location_sp <- as(test_location_sf, "Spatial")
    st_geometry(test_location_sf) <- NULL # turn into data frame
    centroids <- coordinates(test_location_sp)
    test_location_point <- SpatialPointsDataFrame(coords = centroids, data = test_location_sf)
    test_location_sf <- st_as_sf(test_location_point)
    st_crs(test_location_sf) <- st_crs(sampled_data)
    
    # get test data for all years
    test_sgbp = st_intersects(x = sampled_data, y = test_location_sf)
    test_logical = lengths(test_sgbp) > 0
    test_data = sampled_data[test_logical, ]
    
    # get train data for all years 
    train_data <- sampled_data[test_location_sf, , op = st_disjoint]
    
    model = lm(formula, data=train_data) 
    model_intercept = lm(plant ~ 1, data=train_data)
      
    # extract onset coef and R2 from model
    onset_coef <- model$coefficients['onset']
    train_data$residuals <- residuals(model)
    SST <- sum((train_data$plant - mean(train_data$plant))^2)
    SSE <- sum((train_data$residuals - mean(train_data$residuals))^2)
    R2 <- 1 - SSE/SST
    
    # prediction
    prediction <- predict(model, test_data)
    error <- rmse(prediction, test_data$plant)
    
    prediction_intercept <- predict(model_intercept, test_data)
    error_intercept <- rmse(prediction_intercept, test_data$plant)
    
    diff_error <- error_intercept - error
    
    # output
    output <- c(onset_coef, R2, error, diff_error)
    names(output) <- c("onset", "R2", "RMSE", "RMSE_improvement")
    
    prediction_results <- rbind(prediction_results, output)
    
    # plot train and test data for a specific year
  
    # test_year <- test_data[test_data$year == year_oi & test_data$intensity == "DC", ]
    # ggplot(test_year) +
    #    geom_sf(aes(fill = plant)) +
    #    scale_fill_viridis() +
    #    ggtitle(paste("test data for ", year_oi)) +
    #    geom_polygon(data = MT_outline, aes(x = long, y = lat), color = "black", alpha = 0, linetype = 1) +
    #    theme_bw()
    # 
    # train_year <- train_data[train_data$year == year_oi & train_data$intensity == "DC", ]
    # ggplot(train_year) +
    #    geom_sf(aes(fill = plant)) +
    #    scale_fill_viridis() +
    #    ggtitle(paste("train data for ", year_oi)) +
    #    geom_polygon(data = MT_outline, aes(x = long, y = lat), color = "black", alpha = 0, linetype = 1) +
    #    theme_bw()
  }
  
    names(prediction_results) <- names(output)
    prediction_results$index <- 1:nrow(prediction_results)
    
    if (plot_model_evals) {
      plot(prediction_results$index, prediction_results$RMSE, 
           ylab = "RMSE", xlab = "location index", ylim = c(5, 25), type = "l", main = "Prediction RMSE, spatial sampling")
      
      plot(prediction_results$index, prediction_results$RMSE_improvement, 
           ylab = "RMSE", xlab = "location index", ylim = c(0, 10), type = "l", main = "Prediction RMSE improvement over intercept, spatial sampling")
      
      plot(prediction_results$index, prediction_results$onset, type = "l", col = "red", ylab = "coef or R2", xlab = "location index", ylim = c(0.1, 0.7), main = "Prediction onset and R2, spatial sampling")
      lines(prediction_results$index, prediction_results$R2, col = "blue")
      legend(1, 0.65, legend = c("onset coef", "R2"), col = c("red", "blue"), lty= c(1,1))
    }
    
    # map of prediction error
    
    data_in_year$prediction_rmse <- prediction_results$RMSE
    data_in_year_DC <- data_in_year[data_in_year$intensity == "DC", ]
    
    if (plot_model_evals) {
      error_map <- ggplot(data_in_year_DC) +
          geom_sf(aes(fill = prediction_rmse)) +
          scale_fill_viridis() +
          ggtitle("prediction rmse") +
          geom_polygon(data = MT_outline, aes(x = long, y = lat), color = "black", alpha = 0, linetype = 1) +
          theme_bw()
      print(error_map)
    }
    
    return(prediction_results)
    
}

# runs the model with options to eliminate a year
run_model_for_prediction_elimyear <- function(full_data, year_to_elim, elim_year, formula) {
  
  # separate training and test data
  if (elim_year) {
    train_data <- full_data[full_data$year != year_to_elim,]
    test_data <- full_data[full_data$year == year_to_elim,]
  }
  else {
    train_data <- full_data
    test_data <- full_data
  }
  
  model = lm(formula, data=train_data) 
  model_intercept = lm(plant ~ 1, data=train_data)
    
  # extract onset coef and R2 from model
  onset_coef <- model$coefficients['onset']
  train_data$residuals <- residuals(model)
  SST <- sum((train_data$plant - mean(train_data$plant))^2)
  SSE <- sum((train_data$residuals - mean(train_data$residuals))^2)
  R2 <- 1 - SSE/SST
  
  # prediction
  prediction <- predict(model, test_data)
  error <- rmse(prediction, test_data$plant)
  
  prediction_intercept <- predict(model_intercept, test_data)
  error_intercept <- rmse(prediction_intercept, test_data$plant)
  
  diff_error <- error_intercept - error
  
  # output
  output <- c(as.integer(year_to_elim), onset_coef, R2, error, diff_error)
  names(output) <- c("eliminated_year", "onset", "R2", "RMSE", "RMSE_improvement")
  return(output)
}

run_OLS <- function(full_data, predictors, y.var,
                    grid_size, lat_offset, lon_offset, agg_scheme, 
                    plot_samples, plot_model_evals, year_oi) {
  
  # get sampled data
  sampled_data <- get_sampled_data(full_data = cell_sf_tidy, grid_size = grid_size, 
                                   lat_offset = lat_offset, lon_offset = lon_offset, agg_scheme = agg_scheme, 
                                   plot_samples = plot_samples, year_oi = year_oi)
  
  # map of the input data used (for specific year and intensity)
  if (plot_samples) {
    print(plot_cell_onset(year_oi, sampled_data))
    print(plot_cell_plant(year_oi, sampled_data, "DC"))
  }

  # define the formula
  formula.string <- paste(y.var, paste(predictors, collapse = " + "), sep = " ~ ")
  
  f <- as.formula(formula.string)

  # do the model
  
  model = lm(f, data=sampled_data) 
  sampled_data$residuals <- residuals(model)
  sampled_data$fitted.values <- fitted.values(model)
  
  # map of residuals
  if (plot_model_evals) {print(plot_cell_residuals(year_oi, sampled_data, "DC"))}
  
  # evaluation
  
  # 1. coefficients and their p values
  coefficients <- model$coefficients
  onset_coef <- coefficients["onset"]
  intensity_coef <- coefficients["intensitySC"]
  lat_coef <- coefficients["lat"]
  
  p_vals <- summary(model)$coefficients[, ncol(summary(model)$coefficients)]
  onset_pval <- p_vals["onset"]
  intensity_pval <- p_vals["intensitySC"]
  lat_pval <- p_vals["lat"]
  
  # 2. percent data used
  percent_data_used <- 100*nrow(sampled_data)/nrow(full_data)
  
  # 3. R2
  SST <- sum((sampled_data$plant - mean(sampled_data$plant))^2)
  SSE <- sum((sampled_data$residuals - mean(sampled_data$residuals))^2)
  R2 <- 1 - SSE/SST
  
  # 4. exogeneity plots (residual vs predictors)
  if (plot_model_evals) {
    plot(sampled_data$onset, sampled_data$residuals, main = "onset vs residual (exogeneity)")
    plot(sampled_data$lat, sampled_data$residuals, main = "latitude vs residual (exogeneity)")

    plot(sampled_data$year, sampled_data$residuals, main = "latitude vs residual (exogeneity)") 
    abline(h = 0)
  }
  
  # 5. spatial autocorrelation
  
  # need 'one layer': one year, one intensity, set up weights
  to_autocorrelation <- sampled_data[sampled_data$year == year_oi & sampled_data$intensity == "DC", ]
  to_autocorrelation_sp <- as(to_autocorrelation, "Spatial")
  st_geometry(to_autocorrelation) <- NULL # turn to_autocorrelation into data frame
  centroids <- coordinates(to_autocorrelation_sp)
  to_autocorrelation_points <- SpatialPointsDataFrame(coords = centroids, data = to_autocorrelation)
  nb<-knn2nb(knearneigh(to_autocorrelation_points)) 
  lw <- nb2listw(nb, zero.policy = TRUE)
  
  # calculate spatial autocorrelation
  moran_residual <- moran.mc(to_autocorrelation_points$residuals, lw, 500, zero.policy = TRUE)
  moran_onset <- moran.mc(to_autocorrelation_points$onset, lw, 500, zero.policy = TRUE)
  moran_plant <- moran.mc(to_autocorrelation_points$plant, lw, 500, zero.policy = TRUE)
  
  residual_moran <- moran_residual$statistic
  residual_moran_pval <- moran_residual$p.value
  onset_moran <- moran_onset$statistic
  onset_moran_pval <- moran_onset$p.value
  plant_moran <- moran_plant$statistic
  plant_moran_pval <- moran_plant$p.value
  
  # 6. residual plots (homoscedasticity, normality of error, residual-fitted value correlation)
  if (plot_model_evals) {
    test_plots(model, paste("grid size", grid_size, "lat offset", lat_offset, 
                            "lon_offset", lon_offset, "aggregated", agg_scheme))
    plot(sampled_data$plant, sampled_data$fitted.values, 
         main = "fitted value vs actual value of plant", ylab = "fitted.values")
    abline(h = mean(sampled_data$fitted.values), col = "blue")
    abline(v = mean(sampled_data$plant), col = "blue")
    abline(0,1, col = "gray", lwd = 3)
  }
  
  # 7. multicollinearity: look at max absolute correlation between predictors
  predictors <- sampled_data[,c("lat", "onset", "year")]
  st_geometry(predictors) <- NULL
  correlations <- cor(predictors)
  diag(correlations) <- 0 # place 1 with 0 on diagonal
  highest_predictor_correlation <- max(abs(correlations))

  # 8. predictive ability, when eliminate one year
  
  prediction_results_elimyear <- data.frame()

  for (year in 2004:2014) {
    
    result <- run_model_for_prediction_elimyear(sampled_data, year, TRUE, f)
    prediction_results_elimyear <- rbind(prediction_results_elimyear, result)
  }
  
  names(prediction_results_elimyear) <- names(result)
  
  if (plot_model_evals) {
    plot(prediction_results_elimyear$eliminated_year, prediction_results_elimyear$RMSE, 
         ylab = "RMSE", xlab = "eliminated year", ylim = c(5, 25), type = "l", 
         main = "Prediction RMSE, spatial sampling")
    
    plot(prediction_results_elimyear$eliminated_year, prediction_results_elimyear$RMSE_improvement, 
         ylab = "RMSE", xlab = "eliminated year", ylim = c(0, 10), type = "l", 
         main = "Prediction RMSE improvement over intercept, spatial sampling")
    
    plot(prediction_results_elimyear$eliminated_year, prediction_results_elimyear$onset, type = "l", col = "red", 
         ylab = "coef or R2", xlab = "eliminated year", ylim = c(0.1, 0.7), 
         main = "Prediction onset and R2, spatial sampling")
    lines(prediction_results_elimyear$eliminated_year, prediction_results_elimyear$R2, col = "blue")
    legend(2004, 0.65, legend = c("onset coef", "R2"), col = c("red", "blue"), lty= c(1,1))
  }
  
  # 9. predictive ability, location cross validation
  
  prediction_results_elimlocation <- run_model_for_prediction_elimlocation(full_data = sampled_data, 
                                      formula = f, 
                                      year_for_location = 2004, year_oi = 2011,
                                      plot_model_evals = plot_model_evals)
  
  output_list <- list(onset_coef = onset_coef, 
                      intensity_coef = intensity_coef,
                      lat_coef = lat_coef,
                      onset_pval = onset_pval,
                      intensity_pval = intensity_pval,
                      lat_pval = lat_pval,
                      percent_data_used = percent_data_used,
                      R2 = R2,
                      residual_moran = residual_moran,
                      residual_moran_pval = residual_moran_pval,
                      onset_moran = onset_moran,
                      onset_moran_pval = onset_moran_pval,
                      plant_moran = plant_moran,
                      plant_moran_pval = plant_moran_pval,
                      highest_predictor_correlation = highest_predictor_correlation,
                      prediction_results_elimyear = prediction_results_elimyear,
                      prediction_results_elimlocation = prediction_results_elimlocation)
  

  return(output_list)
}

# run one set of spatial sampling/predictor set
model_output <- run_OLS(full_data = cell_sf_tidy, 
                        predictors = c("onset", "intensity",
                          "lat", "lon", # for spatial effects
                          "year_index", # for time effects
                          #"onset:lat", "onset:lon", # interactions: for spatial effects
                          #"onset:year_factor", # interactions: for time effects
                          "onset:intensity" # interactions: for intensity effects
                          ), 
                    y.var = "plant",
                    grid_size = 1, lat_offset = 0.5, lon_offset = 0.5, agg_scheme = TRUE, 
                    plot_samples = TRUE, plot_model_evals = TRUE, 
                    year_oi = 2005)

print(model_output)


```

## functions to evaluate different sampling strategies and predictors

```{r}

# test different grid sizes and aggregation strategies. don't do any offset

results <- data.frame()

for (grid_size in seq(0.5, 1.5, 0.25)) { # 0.25 to 1.75
  for (agg_scheme in c(TRUE, FALSE)) {
    for (lat_offset in seq(0, 1.5, by = 0.25)) { # 0 to 1.5
      for (lon_offset in seq(0, 1.5, by = 0.25)) { # 0 to 1.5
        
        model_output <- run_OLS(full_data = cell_sf_tidy, 
                              predictors = c("onset", "intensity",
                                "lat", "lon", # for spatial effects
                                "year_index", # for time effects
                                "onset:intensity" # interactions: for intensity effects
                                ), 
                          y.var = "plant",
                          grid_size = grid_size, 
                          lat_offset = lat_offset, lon_offset = lon_offset, agg_scheme = agg_scheme, 
                          plot_samples = FALSE, plot_model_evals = FALSE, 
                          year_oi = 2005)
        
        #if (agg_scheme) {agg_scheme_str <- "Agg"}
        #else {agg_scheme_str <- "NotAgg"}
        
        print(paste("grid size", grid_size))
        print(paste("agg scheme", agg_scheme))
    
        selected_output <- as.numeric(c(grid_size, agg_scheme, lat_offset, lon_offset,
                             model_output$onset_coef, model_output$intensity_coef, model_output$lat_coef,
                             model_output$R2, model_output$percent_data_used, 
                             model_output$residual_moran_pval, model_output$onset_moran_pval, 
                             model_output$plant_moran_pval,
                             mean(model_output$prediction_results_elimyear$RMSE, na.rm = TRUE), 
                             mean(model_output$prediction_results_elimlocation$RMSE, na.rm = TRUE)
                             ))
        
        print(selected_output)
        results <- rbind(results, selected_output)
      }
    }
  }
}

names(results) <- c("grid_size", "agg_scheme", "lat_offset", "lon_offset",
                    "onset_coef", "intensity_coef", "lat_coef",
                    "R2", "percent_used", "residual_moran_pval", "onset_moran_pval",
                    "plant_moran_pval",
                    "mean_RMSE_elimyear", "mean_RMSE_elimlocation")
results$agg_scheme <- as.factor(results$agg_scheme)

# SAVED RESULTS
saveRDS(results, "spatial_sampling_results.rds")

# calculate mean and sd for each grid size x agg scheme, for different grid offsets
results_summary <- results %>%
                        group_by(grid_size, agg_scheme) %>%
                        summarize(
                          onset_coef_mean = mean(onset_coef),
                          onset_coef_sd = sd(onset_coef),
                          intensity_coef_mean = mean(intensity_coef),
                          intensity_coef_sd = sd(intensity_coef),
                          lat_coef_mean = mean(lat_coef),
                          lat_coef_sd = sd(lat_coef),
                          R2_mean = mean(R2),
                          R2_sd = sd(R2),
                          percent_used_mean = mean(percent_used),
                          percent_used_sd = sd(percent_used),
                          residual_moran_pval_mean = mean(residual_moran_pval),
                          residual_moran_pval_sd = sd(residual_moran_pval),
                          onset_moran_pval_mean = mean(onset_moran_pval),
                          onset_moran_pval_sd = sd(onset_moran_pval),
                          plant_moran_pval_mean = mean(plant_moran_pval),
                          plant_moran_pval_sd = sd(plant_moran_pval),
                          RMSE_elimyear_mean = mean(mean_RMSE_elimyear),
                          RMSE_elimyear_sd = sd(mean_RMSE_elimyear),
                          RMSE_elimlocation_mean = mean(mean_RMSE_elimlocation),
                          RMSE_elimlocation_sd = sd(mean_RMSE_elimlocation)
                        )

# compare grid sizes and aggregation schemes

plot_results <- function(data, x_data, y_data, x_label, y_label, col_data) {
  ggplot(data, aes(x = x_data, y = y_data, col = col_data)) +
    geom_line() +
    ggtitle(y_label) +
    xlab(x_label) +
    ylab(y_label)+
    labs(col = "Aggregation scheme") +
    theme_bw()
}

plot_results_summary <- function(data, x_data, y_data, y_sd, x_label, y_label, col_data) {
  ggplot(data, aes(x = x_data, y = y_data, col = col_data)) +
    geom_line() +
    geom_errorbar(aes(ymin=y_data-y_sd, ymax=y_data+y_sd), position=position_dodge(0.05)) +
    ggtitle(y_label) +
    xlab(x_label) +
    ylab(y_label)+
    labs(col = "Aggregation scheme") +
    theme_bw()
}

# plot results, summarizing for different grid offsets
plot_results_summary(data = results_summary, 
                     x_data = results_summary$grid_size, y_data = results_summary$onset_coef_mean,
                     y_sd = results_summary$onset_coef_sd,
                    x_label = "grid_size", y_label = "onset coefficient", 
                    col_data = results_summary$agg_scheme)

plot_results_summary(data = results_summary, 
                     x_data = results_summary$grid_size, y_data = results_summary$intensity_coef_mean,
                     y_sd = results_summary$intensity_coef_sd,
                    x_label = "grid_size", y_label = "intensity coefficient", 
                    col_data = results_summary$agg_scheme)

plot_results_summary(data = results_summary, 
                     x_data = results_summary$grid_size, y_data = results_summary$lat_coef_mean,
                     y_sd = results_summary$lat_coef_sd,
                    x_label = "grid_size", y_label = "latitude coefficient", 
                    col_data = results_summary$agg_scheme)

plot_results_summary(data = results_summary, 
                     x_data = results_summary$grid_size, y_data = results_summary$R2_mean,
                     y_sd = results_summary$R2_sd,
                    x_label = "grid_size", y_label = "R2", 
                    col_data = results_summary$agg_scheme)

plot_results_summary(data = results_summary, 
                     x_data = results_summary$grid_size, y_data = results_summary$percent_used_mean,
                     y_sd = results_summary$percent_used_sd,
                    x_label = "grid_size", y_label = "percent used", 
                    col_data = results_summary$agg_scheme)

plot_results_summary(data = results_summary, 
                     x_data = results_summary$grid_size, y_data = results_summary$residual_moran_pval_mean,
                     y_sd = results_summary$residual_moran_pval_sd,
                    x_label = "grid_size", y_label = "residual moran's I p value", 
                    col_data = results_summary$agg_scheme)

plot_results_summary(data = results_summary, 
                     x_data = results_summary$grid_size, y_data = results_summary$onset_moran_pval_mean,
                     y_sd = results_summary$onset_moran_pval_sd,
                    x_label = "grid_size", y_label = "onset moran's I p value", 
                    col_data = results_summary$agg_scheme)

plot_results_summary(data = results_summary, 
                     x_data = results_summary$grid_size, y_data = results_summary$plant_moran_pval_mean,
                     y_sd = results_summary$plant_moran_pval_sd,
                    x_label = "grid_size", y_label = "plant moran's I p value", 
                    col_data = results_summary$agg_scheme)

plot_results_summary(data = results_summary, 
                     x_data = results_summary$grid_size, y_data = results_summary$RMSE_elimlocation_mean,
                     y_sd = results_summary$RMSE_elimlocation_sd,
                    x_label = "grid_size", y_label = "RMSE from eliminating location", 
                    col_data = results_summary$agg_scheme)

plot_results_summary(data = results_summary, 
                     x_data = results_summary$grid_size, y_data = results_summary$RMSE_elimyear_mean,
                     y_sd = results_summary$RMSE_elimyear_sd,
                    x_label = "grid_size", y_label = "RMSE from eliminating year", 
                    col_data = results_summary$agg_scheme)

# plot results for a single grid offset
# plot_results(results, results$grid_size, results$onset_coef, 
#              "grid_size", "onset coefficient", results$agg_scheme)
# 
# plot_results(results, results$grid_size, results$intensity_coef, 
#              "grid_size", "intensitySC coefficient", results$agg_scheme)
# 
# plot_results(results, results$grid_size, results$lat_coef, 
#              "grid_size", "latitude coefficient", results$agg_scheme)
# 
# plot_results(results, results$grid_size, results$R2, 
#              "grid_size", "R2", results$agg_scheme)
# 
# plot_results(results, results$grid_size, results$percent_used, 
#              "grid_size", "percent used", results$agg_scheme)
# 
# plot_results(results, results$grid_size, results$residual_moran_pval, 
#              "grid_size", "p value of residual autocorrelation", results$agg_scheme)
# 
# plot_results(results, results$grid_size, results$onset_moran_pval, 
#              "grid_size", "p value of onset autocorrelation", results$agg_scheme)
# 
# plot_results(results, results$grid_size, results$plant_moran_pval, 
#              "grid_size", "p value of plant autocorrelation", results$agg_scheme)
# 
# plot_results(results, results$grid_size, results$mean_RMSE_elimyear, 
#              "grid_size", "mean prediction RMSE when eliminating a year", results$agg_scheme)
# 
# plot_results(results, results$grid_size, results$mean_RMSE_elimlocation, 
#              "grid_size", "mean prediction RMSE when eliminating a location", results$agg_scheme)

```